

# Full CPTAC study 

```{r, warning=FALSE, message=FALSE}
library(tidyverse)
library(limma)
library(QFeatures)
library(plotly)
library(gridExtra)
library(msqrob2)
library(here)
```
## Read data 

```{r}
pe <- readRDS(file = here("Data/CPTAC_data/CPTAC_QFeatures.RDS"))
```

## Design

- We update the colData with information on the design

```{r}
colData(pe)$lab <- rep(rep(paste0("lab",1:3),each=3),5) %>% as.factor
colData(pe)$condition <- pe[["peptideRaw"]] %>% colnames %>% substr(12,12) %>% as.factor
colData(pe)$spikeConcentration <- rep(c(A = 0.25, B = 0.74, C = 2.22, D = 6.67, E = 20),each = 9)
```

## Preprocessing

### Log-transform

- We calculate how many non zero intensities we have for each peptide and this can be useful for filtering.

```{r}
rowData(pe[["peptideRaw"]])$nNonZero <- rowSums(assay(pe[["peptideRaw"]]) > 0)
```


- Peptides with zero intensities are missing peptides and should be represent
with a `NA` value rather than `0`.

```{r}
pe <- zeroIsNA(pe, "peptideRaw") # convert 0 to NA
```

- Logtransform data with base 2

```{r}
pe <- logTransform(pe, base = 2, i = "peptideRaw", name = "peptideLog")
```


### Filtering

1. Handling overlapping protein groups

In our approach a peptide can map to multiple proteins, as long as there is
none of these proteins present in a smaller subgroup.

```{r}
pe <- filterFeatures(pe, ~ Proteins %in%
                       smallestUniqueGroups(rowData(pe[["peptideLog"]])$Proteins))
```

2. Remove reverse sequences (decoys) and contaminants

We now remove the contaminants, peptides that map to decoy sequences, and proteins
which were only identified by peptides with modifications.

```{r}
pe <- filterFeatures(pe,~Reverse != "+")
pe <- filterFeatures(pe,~ Potential.contaminant != "+")
```

3. Drop peptides that were only identified in one sample

We keep peptides that were observed at last twice in 10% of the samples.

```{r}
pe <- filterNA(pe, i = "peptideLog", pNA = 0.9)
nrow(pe[["peptideLog"]])
```

## Normalization 

```{r}
pe <- normalize(pe, 
                i = "peptideLog", 
                name = "peptideNorm", 
                method = "center.median")
```


## Full dataset

```{r}
set.seed(100)
sce <- pe[["peptideNorm"]]
names(assays(sce)) <- "logintensities"
```


```{r}
saveRDS(object = sce, file = here("Data","CPTAC_data","CPTAC_sce.RDS"))
write.csv(x = assay(sce,"logintensities"), file = here("Data","CPTAC_data","CPTAC_sce.csv"))
```


## Imputation of NAs

```{r}
source(here("R-functions", "Impute_MAR_MNAR.R"))
```

```{r}
set.seed(100)

for(i in c(1:10)){
  MISTOT10_MNAR25 <- imputation_NAs(data = assay(sce, "logintensities"), MISTOT = 0.1, MNAR = 0.25)
  MISTOT10_MNAR50 <- imputation_NAs(data = assay(sce, "logintensities"), MISTOT = 0.1, MNAR = 0.50)
  MISTOT10_MNAR75 <- imputation_NAs(data = assay(sce, "logintensities"), MISTOT = 0.1, MNAR = 0.75)

  
  saveRDS(MISTOT10_MNAR25, paste0(here("Data","CPTAC_data"), "/CPTAC_MISTOT10_MNAR25_",i,".RDS"))
  saveRDS(MISTOT10_MNAR50, paste0(here("Data","CPTAC_data"), "/CPTAC_MISTOT10_MNAR50_",i,".RDS"))
  saveRDS(MISTOT10_MNAR75, paste0(here("Data","CPTAC_data"), "/CPTAC_MISTOT10_MNAR75_",i,".RDS"))

  write.csv(x = MISTOT10_MNAR25$data[rowMeans(is.na(MISTOT10_MNAR25$data)) < 1,], 
            file = paste0(here("Data","CPTAC_data"), "/CPTAC_MISTOT10_MNAR25_",i,".csv"))
  write.csv(x = MISTOT10_MNAR50$data[rowMeans(is.na(MISTOT10_MNAR50$data)) < 1,], 
            file = paste0(here("Data","CPTAC_data"), "/CPTAC_MISTOT10_MNAR50_",i,".csv"))
  write.csv(x = MISTOT10_MNAR75$data[rowMeans(is.na(MISTOT10_MNAR75$data)) < 1,], 
            file = paste0(here("Data","CPTAC_data"), "/CPTAC_MISTOT10_MNAR75_",i,".csv"))
}
```
