---
title: "Leduc Data processing"
output: html_document
date: '2024-03-13'
---


```{r}
library("scpdata")
library(here)
library(MsCoreUtils)
library(ggplot2)
library("scp")

files.sources = list.files(here("R-functions","scp-scplainer","R"))
sapply(paste0(here("R-functions","scp-scplainer", "R",files.sources)), source)

```


```{r}
set.seed(100)
leduc <- leduc2022_pSCoPE()
leduc <- QFeatures(list("peptides" = leduc[["peptides"]]), colData = colData(leduc))
```

## Explore missing values

```{r}
cd <- colData(getWithColData(leduc, "peptides"))
reportMissingValues(leduc, "peptides", by = cd$SampleType)
```

```{r}
ji <- jaccardIndex(leduc, "peptides", by = cd$SampleType)
pji <- ggplot(ji) +
    aes(x = jaccard) +
    geom_histogram() +
    facet_grid(~ by)
```

```{r}
csc <- cumulativeSensitivityCurve(
    leduc, "peptides", by = cd$SampleType, batch = cd$Set,
    niters = 10, nsteps = 30
)
plCSC <- ggplot(csc) +
        aes(x = SampleSize, y = Sensitivity, colour = by) +
        geom_point(size = 1) +
        xlab("# TMT runs included") +
        ggtitle("Cumulative sensitivity curve")

```


```{r}
leduc <- countUniqueFeatures(
    leduc, i = "peptides",
    colDataName = "NumberPeptides"
)


MedianIntensity <- lapply(experiments(leduc), function(x) {
    out <- colMedians(log(assay(x)), na.rm = TRUE)
    names(out) <- colnames(x)
    out
})
names(MedianIntensity) <- NULL
MedianIntensity <- unlist(MedianIntensity)
colData(leduc)[names(MedianIntensity), "MedianIntensity"] <- MedianIntensity

leduc <- medianCVperCell(
    leduc, i = "peptides", groupBy = "Leading.razor.protein.symbol",
    nobs = 3, na.rm = TRUE, colDataName = "MedianCV", norm = "SCoPE2"
)
```


```{r}
ggplot(colData(leduc)) +
    aes(
        y = MedianIntensity,
        x = NumberPeptides,
        color = MedianCV,
        shape = lcbatch
    ) +
    geom_point(size = 2) +
    scale_color_continuous(type = "viridis")



ggplot(data.frame(colData(leduc)[colData(leduc)$NumberPeptides>750, ])) +
    aes(
        y = MedianIntensity,
        x = NumberPeptides,
        color = MedianCV,
        shape = lcbatch
    ) +
    geom_point(size = 2) +
    scale_color_continuous(type = "viridis")
```
```{r}
passQC <- colData(leduc)$NumberPeptides>=750 & colData(leduc)$MedianCV < 0.5
leduc <- subsetByColData(leduc, passQC)

```
## Data processing

Processing steps already performed: 

1. PSM QC: FDR, PIF, contaminants and decoy, SCR, cell total intensity.
2. Reference normalization
3. Remove duplicate PSM per peptide
4. Zero is NA
5. Cell QC based on CV

For the remainder of this experiment, I keep peptides that have 
at most 90% missing values, meaning they are observed in at least 10%
of the cells. This removes a lot of peptides...

```{r}
nrows(leduc)[["peptides"]]
leduc <- filterNA(leduc, i = "peptides", pNA = 0.9)
nrows(leduc)[["peptides"]]
```

I also perform the following steps, some that I will compare against 
each other

1. Log2 transformation
2. Median centering the cells

```{r}
## 1.
colData(leduc[["peptides"]]) <- NULL
leduc <- logTransform(leduc, "peptides", name = "peptides_log2")
## 2.
leduc <- scp:::normalizeSCP(
    leduc, i = "peptides_log2", method = "center.median",
    name = "peptides_norm_samples"
)
```


```{r}
peptides <- getWithColData(leduc, "peptides_norm_samples")
names(assays(peptides)) <- "logintensities"
```


```{r}
scprotein_assay <- assay(peptides,"logintensities")
scprotein_assay <- (cbind(rowData(peptides)$Leading.razor.protein.symbol,rownames(scprotein_assay), scprotein_assay))
colnames(scprotein_assay)[1:2] <- c("protein","peptide")
rownames(scprotein_assay) <- NULL
scprotein_assay <- as.data.frame(scprotein_assay)
scprotein_assay[,3:ncol(scprotein_assay)] <- lapply(scprotein_assay[,3:ncol(scprotein_assay)], as.numeric)
write.csv(x = scprotein_assay, file = here("Analyses/Leduc/scProtein/data/Peptides-raw-train1.csv"), row.names = F)

scprotein_assay <- scprotein_assay[,-1]
colnames(scprotein_assay)[1] <- "protein"

Cells <- colData(peptides)$SampleType
Cells[colData(peptides)$MelanomaSubCluster == "A"] <- "Melanoma_A"
Cells[colData(peptides)$MelanomaSubCluster == "B"] <- "Melanoma_B"
Batch <- colData(peptides)$Set
Batch_chromatocraphy <- colData(peptides)$Channel
cells_df <- rbind(Cells, Batch, Batch_chromatocraphy)
colnames(cells_df) <- colnames(peptides)
rownames(cells_df) <- c("celltype", "batch_sort", "batch_chromatography")
cells_df[1:3,1:5]
```




```{r}
saveRDS(object = peptides, file = here("Data","leduc2022_pSCoPE_data","leduc_sce.RDS"))
write.csv(x = assay(peptides,"logintensities"), 
          file = here("Data","leduc2022_pSCoPE_data","leduc_sce.csv"))

write.csv(x = scprotein_assay, file = here("Analyses/Leduc/scProtein/data/Peptides-raw.csv"), row.names = F)
write.csv(x = cells_df, file = here("Analyses/Leduc/scProtein/data/Cells.csv"))

```


## Imputation of NAs

```{r}
source(here("R-functions", "Impute_MAR_MNAR.R"))
```




```{r}

set.seed(100)
for(i in c(1:10)){
  

MISTOT10_MNAR25 <- imputation_NAs(data = assay(peptides, "logintensities"), MISTOT = 0.1, MNAR = 0.25)
MISTOT10_MNAR50 <- imputation_NAs(data = assay(peptides, "logintensities"), MISTOT = 0.1, MNAR = 0.50)
MISTOT10_MNAR75 <- imputation_NAs(data = assay(peptides, "logintensities"), MISTOT = 0.1, MNAR = 0.75)

saveRDS(MISTOT10_MNAR25,paste0(here("Data","leduc2022_pSCoPE_data"),"/leduc_MISTOT10_MNAR25_", i, ".RDS"))
saveRDS(MISTOT10_MNAR50, paste0(here("Data","leduc2022_pSCoPE_data"),"/leduc_MISTOT10_MNAR50_", i, ".RDS"))
saveRDS(MISTOT10_MNAR75, paste0(here("Data","leduc2022_pSCoPE_data"),"/leduc_MISTOT10_MNAR75_", i, ".RDS"))

write.csv(x = MISTOT10_MNAR25$data, file = paste0(here("Data","leduc2022_pSCoPE_data"),"/leduc_MISTOT10_MNAR25_", i, ".csv"))
write.csv(x = MISTOT10_MNAR50$data, file = paste0(here("Data","leduc2022_pSCoPE_data"),"/leduc_MISTOT10_MNAR50_", i, ".csv"))
write.csv(x = MISTOT10_MNAR75$data, file = paste0(here("Data","leduc2022_pSCoPE_data"),"/leduc_MISTOT10_MNAR75_", i, ".csv"))

}
```
